Unemployment Effects: How Multi-Year Periods of Unemployment Influence Feelings Towards the Government

### 2020-05-08 analysis.Rmd
added some stuff to my literature review

```{r extension data}

@@

# SHP
# I run six regressions
# Two for each question (trust in government, satisfaction with democracy, and life satisfaction)
# The first regression is just with unemployed as the indenpendent variable
# The second regression includes the controls (age, education, and member)
# The regression uses a pooling technique (plm) to track observations across units (households) 
# over time

@@

```{r stargazer data}
wfe_regression_data <- data.frame(
  c1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,1]),
  c2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,1]),
  c3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,1]),
  c4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,1]),
  c5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,1]),
  c6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,1]),
  se1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,2]),
  se2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,2]),
  se3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,2]),
  se4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,2]),
  se5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,2]),
  se6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,2]),
  t1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,3]),
  t2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,3]),
  t3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,3]),
  t4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,3]),
  t5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,3]),
  t6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,3]),
  p1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,4]),
  p2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,4]),
  p3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,4]),
  p4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,4]),
  p5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,4]),
  p6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,4])
) %>%
  insertRows(3:5)
```

@@

>>>>

```{r extension data}
# Create the lagged unemployed variable
# Arrange the panel data by unit and time so the lag command works
# Did this for both the liss and the shp data

@@

# SHP
# I run six regressions
# Two for each question (trust in government, satisfaction with democracy, and life satisfaction)
# The first set of regressions is just with the unemployed and lagged unemployed interaction as the indenpendent variable
# The second set of regressions includes the controls (age, education, and member)
# The regression uses a pooling technique (plm) to track observations across units (households) 
# over time

@@

```{r stargazer data}

# Because I the wfe regression model does not create an output suitable for stargazer, 
# Bauer had to input the data manually

# After I added the interaction term the manual entries didn't match the labels so I
# had to create this new dataframe with added blank rows to get the data to match

wfe_regression_data <- data.frame(
  c1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,1]),
  c2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,1]),
  c3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,1]),
  c4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,1]),
  c5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,1]),
  c6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,1]),
  se1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,2]),
  se2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,2]),
  se3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,2]),
  se4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,2]),
  se5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,2]),
  se6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,2]),
  t1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,3]),
  t2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,3]),
  t3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,3]),
  t4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,3]),
  t5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,3]),
  t6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,3]),
  p1 = as.numeric(summary(shp.fit.wfe.tig.ext)[2]$coefficients[,4]),
  p2 = as.numeric(summary(shp.fit.wfe.swd.ext)[2]$coefficients[,4]),
  p3 = as.numeric(summary(shp.fit.wfe.ls.ext)[2]$coefficients[,4]),
  p4 = as.numeric(summary(liss.fit.wfe.tig.ext)[2]$coefficients[,4]),
  p5 = as.numeric(summary(liss.fit.wfe.swd.ext)[2]$coefficients[,4]),
  p6 = as.numeric(summary(liss.fit.wfe.ls.ext)[2]$coefficients[,4])
) %>%
  insertRows(3:5)
```

@@

# The data from the regressions without controls are pulled from my datatable (created # above)
# The data from the regressions with controls are pulled from the regression outputs

### 2020-05-08 paper.Rmd
added some stuff to my literature review

"Unemployment, Trust in Government, and Satisfaction with Democracy: An Empirical Investigation", authored by Paul Bauer, explores how unemployment affects a person's trust in the government and their views towards democracy in general (@bauer2018). Bauer utilizes panel data from two household surveys, one in Switzerland and one in the Netherlands, to evaluate the relationship between unemployment status and feelings towards government on an individual level. Due to prior findings on the link between life satisfaction and unemployment, Bauer uses that relationship as a base metric to evaluate the magnitude of the other effects upon. Bauer finds that a household's unemployment status has no consistent effect on either trust in the government or satisfaction with democracy. These findings contradict his initial hypothesis that unemployment would lead people to have lower levels of trust and satisfaction with the government and democratic institutions. However, Bauer does find that unemployment is negatively associated with life satisfaction. People who are unemployed are generally less satisfied with their lives than those who are employed. This finding is consistent with previous research on the topic (@hudson2006; @jimeno2007). Moving forward, Bauer indicates an interest in increasing the sample size of the studied group to further explore heterogenous treatment effects across various subgroups of individuals. For example, he hypothesizes that the effect of unemployment on attitudes may be greater among people traditionally disadvantaged in the labor market such as women and workers of a lower-class.

Starting with the replication materials provided by Bauer through the Harvard Dataverse, I worked to replicate the findings of Bauer's paper (@dataverse).^[The replication materials for this paper and my subsequent analyses can be found at my [Github Repo](https://github.com/PGQuinton/gov1006-finalproject-2.0) and the published paper is available through [bookdown](https://bookdown.org/pquinton/milestone7/milestone7.pdf).] As Bauer did, I conducted my replication and subsequent extension in RStudio (@RStudio). Thanks to Bauer's well documented code and use of RStudio, my replication process was relatively straight forward though there were a couple of discrepencies which I will go into in further detail in the "Replication" section of this paper.

@@

The impetus for Bauer's paper relies on the fundamental assumption that should a person blame the government or its institutions for their current situation in life, a decrease in the quality of that situation would result in a corresponding decrease in their satisfaction with the government. This assumption stems from the work of John Hudson who examined the relationship between trust in institutions and one's subjective well-being (@hudson2006). Research shows a statistically significant relationship between one's employment status and their life satisfaction (@jimeno2007). As one moves from employed to unemployed, their life satisfaction tends to decrease. Given the two findings, the aim of Bauer's work is to test the hypothesis that unemployment leads to lower levels of trust in government and satisfaction with democracy.

Bauer's work draws on the theoretical concepts of James Laurence who argued that the loss of one's job may lead to a deterioration of their trust in the economic system (@laurence2015). Furthermore, the loss of one's job brings people into closer contact with bureaucratic institutions and government social programs. Such interactions may incite a frusturation with the way in which the government supports individuals experiencing joblessness and could lead to a dissolusionment with the current government (@waters2007). It is from this theoretical basis that Bauer hypothesizes unemployment to be negatively associated with trust in government and satisfaction with democracy. There is, however, an alternative theory. Should a person feel supported and appreciated by the government and its programs during their period of joblessness, their feelings towards the government may not change or may even improve (@roth2009).

@@

In 2018, Paul Bauer published the paper, "Unemployment, Trust in Government, and Satisfaction with Democracy: An Empirical Investigation" in which he explored the relationship between a person's unemployment status and their trust in government and satisfaction with democracy (@bauer2018). Using data from the SHP and LISS household surveys, Bauer found that there was no statistically significant effect between unemployment and either trust in government or statisfaction with democracy. These findings challenged his hypothesis that short-term unemployment would lead people to lose faith in the government and its institutions. Bauer's hypothesis was based on theory and prior findings indicating a link between unemployment and decreasing life satisfaction.

@@

Both my results, and Bauer's, can offer encouraging interpretations. As Felix Roth theorized, positive interactions with government programs and institutions during a period of unemployment may lead to improvements in that individual's feelings towards the government, or, at the very least, prevent a strong decrease in attitude (@roth2009). Thus, the lack of statistically significant effects in these regression models may indicate highly effective programs for those experiencing joblessness. Looking at the countries surveyed, Switzerland and the Netherlands, such an optimistic interpretation is plausible given the robustness of their social welfare programs. However, the lack of a consistent effect could also stem from limitations with the data. As such, futher studies making use of longer term data across several countries are warranted to more conclusively understand the link between unemployment and feelings towards government.

@@

My replications of the first two tables in Bauer's paper are below.

\input{./replication-table-1.tex}

@@

The actual tables created by Bauer are below. As you can see, there are discrepencies between my Table 2 and his.

\input{./table-1.tex}

>>>>

"Unemployment, Trust in Government, and Satisfaction with Democracy: An Empirical Investigation", authored by Paul Bauer, explores how unemployment affects a person's trust in the government and their views towards democracy in general.^[@bauer2018] Bauer utilizes panel data from two household surveys, one in Switzerland and one in the Netherlands, to evaluate the relationship between unemployment status and feelings towards government on an individual level. Due to prior findings on the link between life satisfaction and unemployment, Bauer uses that relationship as a base metric to evaluate the magnitude of the other effects upon. Bauer finds that a household's unemployment status has no consistent effect on either trust in the government or satisfaction with democracy. These findings contradict the hypothesis that unemployment would lead people to have lower levels of trust and satisfaction with the government and democratic institutions. However, Bauer does find that unemployment is negatively associated with life satisfaction. People who are unemployed are generally less satisfied with their lives than those who are employed. This finding is consistent with previous research on the topic.^[@hudson2006; @jimeno2007] Moving forward, Bauer indicates an interest in increasing the sample size of the studied group to further explore heterogenous treatment effects across various subgroups of individuals. For example, he hypothesizes that the effect of unemployment on attitudes may be greater among people traditionally disadvantaged in the labor market such as women and workers of a lower-class.

Starting with the replication materials provided by Bauer through the Harvard Dataverse, I worked to replicate the findings of Bauer's paper.^[@dataverse] As Bauer did, I conducted my replication and subsequent extension in RStudio.^[@RStudio] Thanks to Bauer's well documented code and use of RStudio, my replication process was relatively straight forward though there were a couple of discrepencies which I will go into in further detail in the "Replication" section of this paper.^[The replication materials for this paper and my subsequent analyses can be found at my [Github Repo](https://github.com/PGQuinton/gov1006-finalproject-2.0) and the published paper is available through [bookdown](https://bookdown.org/pquinton/milestone7/milestone7.pdf).]

@@

The impetus for Bauer's paper relies on the fundamental assumption that should a person blame the government or its institutions for their current situation in life, a decrease in the quality of that situation would result in a corresponding decrease in their satisfaction with the government.^[@bauer2018] This assumption stems from the work of John Hudson who examined the relationship between trust in institutions and one's subjective well-being.^[@hudson2006] Hudson found that unemployed people have,on average, lower levels of trust in government institutions from the federal to the local level.^[@hudson2006] Research shows a statistically significant relationship between one's employment status and their life satisfaction.^[@jimeno2007] As one moves from employed to unemployed, their life satisfaction tends to decrease. Given the two findings, the aim of Bauer's work is to test the hypothesis that unemployment leads to lower levels of trust in government and satisfaction with democracy.

Bauer's work draws on the theoretical concepts of James Laurence who argued that the loss of one's job may lead to a deterioration of their trust in the economic system.^[@laurence2015] The decrease in trust may be born out of blame for the situation however, it may also be due to increased exposure to government programs. The loss of one's job brings people into closer contact with bureaucratic institutions and government social programs.^[@bauer2018] Such interactions may incite a frusturation with the way in which the government supports individuals experiencing joblessness and could lead to a dissolusionment with the current government.^[@waters2007] It is from this theoretical basis that Bauer poses two hypotheses. The first, unemployment is negatively associated with trust in government and satisfaction with democracy.^[@bauer2018] Alternatively, should a person feel supported and appreciated by the government and its programs during their period of joblessness, their feelings towards the government may not change or may even improve.^[@roth2009] Thus, the second, or null, hypothesis is that unemployment will lead to no changes in trust in government or satisfaction with democracy.

@@

In 2018, Paul Bauer published the paper, "Unemployment, Trust in Government, and Satisfaction with Democracy: An Empirical Investigation" in which he explored the relationship between a person's unemployment status and their trust in government and satisfaction with democracy.^[@bauer2018] Using data from the SHP and LISS household surveys, Bauer found that there was no statistically significant effect between unemployment and either trust in government or statisfaction with democracy. These findings challenged the hypothesis that short-term unemployment would lead people to lose faith in the government and its institutions. That hypothesis was based on theory and prior findings indicating a link between unemployment and decreasing life satisfaction.

@@

Both my results, and Bauer's, can offer encouraging interpretations. As Felix Roth theorized, positive interactions with government programs and institutions during a period of unemployment may lead to improvements in that individual's feelings towards the government, or, at the very least, prevent a strong decrease in attitude.^[@roth2009] Thus, the lack of statistically significant effects in these regression models may indicate highly effective programs for those experiencing joblessness. Looking at the countries surveyed, Switzerland and the Netherlands, such an optimistic interpretation is plausible given the robustness of their social welfare programs. However, the lack of a consistent effect could also stem from limitations with the data. As such, futher studies making use of longer term data across several countries are warranted to more conclusively understand the link between unemployment and feelings towards government.

@@

My replication of the second table from Bauer's paper is below.

@@

The actual table created by Bauer are below. As you can see, there are discrepencies between my Table 2 and his.
